---
title: "Week 9 In-Class Assignment"
author: "Felix Ho"
date: "2024-11-4"
format: html
embed-resources: true
---


```{r}
#| include: false
library(here)
library(tidyverse)
library(mice)
library(texreg)
```


```{r}
#| include: false
# Use source() to call the R script that creates a data set.
source(here("R", "merge_alldata.R"))
```


Log transform GDP and use it as one of the predictors. 
```{r}
mergealldata$loggdp <- log(mergealldata$gdp1000)
mergealldata$pctpopdens <- mergealldata$popdens / 100
```


Apply multiple imputation:
```{r}
# Convert ISO to numeric.
midata <- mergealldata |>
  mutate(ISOnum = as.numeric(as.factor(mergealldata$ISO))) |>
  select(-country_name, -ISO)

# Dry run to get meth and pred.
mice0  <- mice(midata, seed = 100, m = 5, maxit = 0, print = F)
```


Edit meth and pred.
```{r}
meth <- mice0$method
meth[c("urban", "male_edu", "temp", "rainfall1000", "matmor", "infmor", "neomor", "un5mor", "loggdp", "pctpopdens")] <- "2l.lmer"

pred <- mice0$predictorMatrix
pred[c("urban", "male_edu", "temp", "rainfall1000", "matmor", "infmor", "neomor", "un5mor", "loggdp", "pctpopdens"), "ISOnum"] <- -2
```


Perform multiple imputation with m = 10 imputations:
```{r}
#| eval: false
# This ran for 11 minutes.
mice.multi.out  <- mice(midata, seed = 100, m = 10, maxit = 20,
                        method = meth,
                        predictorMatrix = pred)

# Save the R object "mice.multi.out" as an RData file with .Rda extension.
save(mice.multi.out, file = here("data", "miceout.Rda"))
```


```{r}
# Check for convergence.
load(here("data", "miceout.Rda"))
plot(mice.multi.out)
```



Fit the four regression models using the same 10 imputed data sets using the with() function. plm() cannot be used within with(), so need to use lm().
```{r}
fit.matmormod <- with(mice.multi.out,
                      matmormod <- lm(matmor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year)))
fit.un5mormod <- with(mice.multi.out,
                      un5mormod <- lm(un5mor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year)))
fit.infmormod <- with(mice.multi.out,
                      infmormod <- lm(infmor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year)))
fit.neomormod <- with(mice.multi.out,
                      neomormod <- lm(neomor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year)))

# summary(pool(fit.matmormod))
# summary(pool(fit.un5mormod))
# summary(pool(fit.infmormod))
# summary(pool(fit.neomormod))

mi.matmormod <- pool(fit.matmormod)
mi.un5mormod <- pool(fit.un5mormod)
mi.infmormod <- pool(fit.infmormod)
mi.neomormod <- pool(fit.neomormod)
```


Complete case analyses:
```{r}
matmormod1 <- lm(matmor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year),
                 data = midata)
un5mormod1 <- lm(un5mor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year),
                 data = midata)
infmormod1 <- lm(infmor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year),
                 data = midata)
neomormod1 <- lm(neomor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + ISOnum + as.factor(year),
                 data = midata)
```


Create a table that shows coefficients and standard errors from complete-case analyses and multiple imputation analysis:
```{r}
screenreg(list(matmormod1, mi.matmormod, un5mormod1, mi.un5mormod, infmormod1, mi.infmormod, neomormod1, mi.neomormod), 
               custom.model.names = c("Maternal mortality ratio (Complete case)",
                            "Maternal mortality ratio (Multiple imputation)",
                            "Under-5 mortality rate (Complete case)", 
                            "Under-5 mortality rate (Multiple imputation)",
                            "Infant mortality rate (Complete case)", 
                            "Infant mortality rate (Multiple imputation)",
                            "Neonatal mortality rate (Complete case)",
                            "Neonatal mortality rate (Multiple imputation)"))
```