---
title: "Week 9 In-Class Assignment"
author: "Felix Ho"
date: "2024-11-4"
format: html
embed-resources: true
---


```{r}
#| label: load-packages
#| include: false
library(here)
library(tidyverse)
library(mice)
library(texreg)
```


```{r}
#| include: false
# Use source() to call the R script that creates a data set.
source(here("R", "merge_alldata.R"))
```


Log transform GDP and use it as one of the predictors. 
```{r}
mergealldata$loggdp <- log(mergealldata$gdp1000)
mergealldata$pctpopdens <- mergealldata$popdens / 100
mergealldata <- mergealldata %>% 
  select(-popdens, -gdp1000)
```


Apply multiple imputation:
```{r}
# Convert ISO to numeric.
midata <- mergealldata |>
  mutate(ISOnum = as.numeric(as.factor(mergealldata$ISO))) |>
  select(-country_name, -ISO, -region, -OECD2023)

# Dry run to get meth and pred.
mice0  <- mice(midata, seed = 100, m = 5, maxit = 0, print = F)
```


Edit meth and pred.
```{r}
meth <- mice0$method
meth[c("urban", "male_edu", "temp", "rainfall1000", "matmor", "infmor", "neomor", 
       "un5mor", "loggdp", "pctpopdens")] <- "2l.lmer"

pred <- mice0$predictorMatrix
pred[c("urban", "male_edu", "temp", "rainfall1000", "matmor", "infmor", "neomor", 
       "un5mor", "loggdp", "pctpopdens"), "ISOnum"] <- -2
```


Perform multiple imputation with m = 10 imputations:
```{r}
#| eval: false
mice.multi.out  <- mice(midata, seed = 100, m = 10, maxit = 20,
                        method = meth,
                        predictorMatrix = pred)

# Save the R object "mice.multi.out" as an RData file with .Rda extension.
save(mice.multi.out, file = here("data", "imputed_data.Rda"))
```


Check for convergence.
```{r}
load(here("data", "imputed_data.Rda"))
plot(mice.multi.out)
```


Fit the four regression models using the same 10 imputed data sets using the with() function. plm() cannot be used within with(), so need to use lm().
```{r}
fit.matmormod <- with(mice.multi.out,
                      lm(matmor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + as.factor(ISOnum) + as.factor(year)))
fit.un5mormod <- with(mice.multi.out,
                      lm(un5mor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + as.factor(ISOnum) + as.factor(year)))
fit.infmormod <- with(mice.multi.out,
                      lm(infmor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + as.factor(ISOnum) + as.factor(year)))
fit.neomormod <- with(mice.multi.out,
                      lm(neomor ~ -1 + armcon + loggdp + OECD + 
                                        pctpopdens + urban + agedep + male_edu + 
                                        temp + rainfall1000 + earthquake + 
                                        drought + as.factor(ISOnum) + as.factor(year)))

# summary(pool(fit.matmormod))
# summary(pool(fit.un5mormod))
# summary(pool(fit.infmormod))
# summary(pool(fit.neomormod))

mi.matmormod <- pool(fit.matmormod)
mi.un5mormod <- pool(fit.un5mormod)
mi.infmormod <- pool(fit.infmormod)
mi.neomormod <- pool(fit.neomormod)
```


Complete case analyses:
```{r}
preds <- as.formula(" ~ -1 + armcon + loggdp + OECD + pctpopdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought + 
                  as.factor(ISO) + as.factor(year)")

cc.matmormod <- lm(update.formula(preds, matmor ~ .), data = mergealldata)
cc.un5mormod <- lm(update.formula(preds, un5mor ~ .), data = mergealldata)
cc.infmormod <- lm(update.formula(preds, infmor ~ .), data = mergealldata)
cc.neomormod <- lm(update.formula(preds, neomor ~ .), data = mergealldata)
```


Create a table that shows coefficients and 95% confidence intervals from complete-case analyses and multiple imputation analyses:
```{r}
tosave <- list(cc.matmormod, mi.matmormod, cc.un5mormod, mi.un5mormod, cc.infmormod, 
               mi.infmormod, cc.neomormod, mi.neomormod)

keepvars <- list("armcon" = "Armed conflict - binary",
                 "loggdp" = "log(GDP)",
                 "OECD" = "OECD",
                 "pctpopdens" = "Population density",
                 "urban" = "Urban residence",
                 "agedep" = "Age dependency ratio",
                 "male_edu" = "Male education",
                 "temp" = "Average temperature",
                 "rainfall1000" = "Average rainfall",
                 "earthquake" = "Earthquake",
                 "drought" = "Drought")

screenreg(tosave, ci.force = TRUE, custom.coef.map = keepvars, 
          custom.model.names = c("Mat CC", "Mat MI", "Un5 CC", "Un5 MI", "Inf CC", 
                                 "Inf MI", "Neo CC", "Neo MI"))

save(tosave, file = here("data", "complete-case_mi_output.Rda"))
```