---
title: "Week 8 In-Class Assignment"
author: "Felix Ho"
date: "2024-10-28"
format: html
embed-resources: true
---

```{r}
#| include: false
library(here)
library(tidyverse)
library(plm)
library(stargazer)
```



```{r}
#| include: false
# Use source() to call the R script that creates a data set.
source(here("R", "merge_alldata.R"))
```



Log transform GDP and use it as one of the predictors. Fit the four mortality models using plm().

```{r}
mergealldata$loggdp <- log(mergealldata$gdp1000)
mergealldata$popdens <- mergealldata$popdens / 100
preds1 <- as.formula(" ~ armcon + loggdp + OECD + popdens + urban + 
                  agedep + male_edu + temp + rainfall1000 + earthquake + drought")

matmormod1 <- plm(update.formula(preds1, matmor ~ .), index = c("ISO", "year"),
                  effect = "twoways", model = "within", data = mergealldata)
un5mormod1 <- plm(update.formula(preds1, un5mor ~ .), index = c("ISO", "year"),
                  effect = "twoways", model = "within", data = mergealldata)
infmormod1 <- plm(update.formula(preds1, infmor ~ .), index = c("ISO", "year"),
                  effect = "twoways", model = "within", data = mergealldata)
neomormod1 <- plm(update.formula(preds1, neomor ~ .), index = c("ISO", "year"),
                  effect = "twoways", model = "within", data = mergealldata)
```



Create a table similar to Table 2 in the paper for only the binary armed conflict specification.
```{r}
stargazer(matmormod1, un5mormod1, infmormod1, neomormod1, type = "text", ci = TRUE, 
          ci.level = 0.95, omit.stat = c("rsq", "f", "ser", "adj.rsq"), 
          column.labels = c("Maternal mortality ratio", 
                            "Under-5 mortality rate", 
                            "Infant mortality rate", 
                            "Neonatal mortality rate"), 
          covariate.labels = c("Armed conflict - binary", "log(GDP per capita)", 
                               "OECD member", "Population density", "Urban residence", 
                               "Age dependency ratio", "Male education", "Temperature", 
                               "Rainfall", "Earthquakes", "Droughts"), 
          title = "Table: The association between armed conflict and maternal and child mortality (adjusted beta coefficients, 95% confidence intervals)", 
          dep.var.labels = " ", align = TRUE, no.space = TRUE, single.row = TRUE)
```
Set align argument to TRUE, so that coefficients in each column are aligned along the decimal point. Remove all empty lines from the table (using no.space). Use omit.stat to leave out several statistics - R2 ("rsq"), F-statistic ("f"), residual standard error ("ser"), and Adjusted R2 ("adj.rsq"). Use column.labels and covariate.labels arguments to label each of the dependent and independent variables. Report 95 percent confidence intervals (using ci and ci.level). Report the coefficients and confidence intervals on the same row (using single.row). Instead of writing omit.stat = c("rsq", "f", "ser", "adj.rsq"), can write keep.stat = "n" to keep only the number of observations. To produce ASCII text output, rather than LATEX code, write type = "text". Can change the order of the independent variables using the order argument. 